An equivalence of compositional frameworks

### 2019-04-23 EquivCompFrameworks_20180911kc.tex
added and edited various needed parts. Next step, the proof of the theorem!

\section*{Various issues-in-progress}
{\chris If Kenny likes, we can use this as a section for all the current issues that need to be resolved, or what we are currently working on or something!}
\begin{itemize}
\item {\chris I think structured cospans should be introduced/recalled way earlier! This is a paper that compares the two,
and both should be presented early.}
\item Sure, I'm fine with that. What section do you think we should bring it up in, or maybe create a new section entirely to mention it?
\end{itemize}

@@

@@

\section{An equivalence of symmetric monoidal double categories} \label{EquivDoubleCats}
In this section we prove that two different frameworks which utilize symmetric monoidal double categories, namely decorated cospans and `structured cospans', are equivalent under the conditions required from the previous section that allow for a left adjoint $L \colon \mathsf{A} \to \mathsf{X}$ to be obtained from a symmetric lax monoidal pseudofunctor $F \colon \mathsf{A} \to \mathbf{Cat}$. The double category version of decorataed cospans is given in Section \ref{DecCospansDoubleCat}. The other framework also utilizing symmetric monoidal double categories due to the first two authors and which goes by the name of `structured cospans', is introduced below. Once again following the notation of Shulman \cite{Shul2}, given a double category $\mathbb{A}$, we write $_f \mathbb{A}_g(M,N)$ for the set of 2-morphisms in $\mathbb{A}$ of the form:
\[
  \xymatrix@-.5pc{
    A \ar[r]|{|}^{M}  \ar[d]_f \ar@{}[dr]|{\Downarrow a}&
    B\ar[d]^g\\
    C \ar[r]|{|}_N & D
  }
\]
We call $M$ and $N$ the \define{horizontal source and target} of the 2-morphism $a$, respectively, and likewise we call $f$ and $g$ the \define{vertical source and target} of the 2-morphism $a$, respectively. Thus $_f \mathbb{A}_g(M,N)$ denotes the set of 2-morphisms in $\mathbb{A}$ with horizontal source and target $M$ and $N$ and vertical source and target $f$ and $g$.
\begin{defn}
A (possibly lax or oplax) double functor $\mathbb{F} \colon \mathbb{A} \to \mathbb{X}$ is \define{full} (respectively, \define{faithful}) if $\mathbb{F}_0 \colon \mathbb{A}_0 \to \mathbb{X}_0$ is full (respectively, faithful) and each map $$\mathbb{F}_1 \colon _f \mathbb{A}_g(M,N) \to _{\mathbb{F}(f)} \mathbb{X}_{\mathbb{F}(g)}(\mathbb{F}(M),\mathbb{F}(N))$$ is surjective (respectively, injective).
\end{defn}
\begin{defn}
A (possibly lax or oplax) double functor $\mathbb{F} \colon \mathbb{A} \to \mathbb{X}$ is \define{essentially surjective} if we can simultaneously make the following choices:
\begin{enumerate}
\item{For each object $x \in \mathbb{X}$, we can find an object $a \in \mathbb{A}$ together with a vertical 1-isomorphism $\alpha_x \colon \mathbb{F}(a) \to x$, and}
\item{For each horizontal 1-cell $N \colon x_1 \tobar x_2$  of $\mathbb{X}$, we can find a horizontal 1-cell $M \colon a_1 \tobar a_2$ of $\mathbb{A}$ and a 2-isomorphism $a_{N}$ of $\mathbb{X}$ as in the following diagram:
\[
  \xymatrix@-.5pc{
    \mathbb{F}(a_1) \ar[r]|{|}^{\mathbb{F}(M)}  \ar[d]_{\alpha_{x_1}} \ar@{}[dr]|{\Downarrow a_N}&
    \mathbb{F}(a_1) \ar[d]^{\alpha_{x_2}}\\
    x_1 \ar[r]|{|}_N & x_2
  }
\]
}
\end{enumerate}
\end{defn}
\begin{defn}
A double functor $\mathbb{F} \colon \mathbb{A} \to \mathbb{X}$ is \define{strong} if the comparison and unit constraints are globular isomorphisms, meaning that for each composable pair of horizontal 1-cells $M$ and $N$ we have a natural isomorphism $$\mathbb{F}_{M,N} \colon \mathbb{F}(M) \odot \mathbb{F}(N) \xrightarrow{\sim} \mathbb{F}(M \odot N)$$and for each object $a \in \mathbb{A}$ a natural isomorphism $$\mathbb{F}_a \colon \hat{U}_{\mathbb{F}(a)} \xrightarrow{\sim} \mathbb{F}(U_a).$$
\end{defn}
\begin{thm}[Shulman,7.8]\label{ShulDubEquiv}
Given a strong double functor $\mathbb{F} \colon \mathbb{A} \to \mathbb{X}$, $\mathbb{F}$ is part of a double equivalence if and only if $\mathbb{F}$ is full, faithful and essentially surjective.
\end{thm}

@@

{\chris Statement of Theorem, and proof, will be heavily edited soon!}
\begin{thm}\label{Equiv}
Let $\mathsf{A}$ and $\mathsf{X}$ be categories with finite colimits and $F \colon \mathsf{A} \to \mathbf{Cat}$ a symmetric lax monoidal pseudofunctor. Let $L \colon \mathsf{A} \to \int{F} \cong \mathsf{X}$ be the left adjoint of the Grothendieck construction $R \colon \int{F} \to \mathsf{A}$ of $F$. Then the symmetric monoidal double category $_L \mathbb{C}\textnormal{sp}(\mathsf{X})$ utilizing structured cospans and the symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$ utilizing decorated cospans are equivalent as symmetric monoidal double categories.
\end{thm}
\begin{proof}
To prove this, we define a double functor $\mathbb{E} \colon _L \mathbb{C}\textnormal{sp}(\mathsf{X}) \to F\mathbb{C}\textnormal{sp}$ as follows: the object component of the double functor $\mathbb{E}$ is given by $\mathbb{E}_0 = \id_{\mathsf{A}}$ as both double categories $_L \mathbb{C}\textnormal{sp}(\mathrm{X})$ and $F\mathbb{C}\textnormal{sp}$ have objects and morphisms of $\mathsf{A}$ as objects and vertical 1-morphisms, respectively. The functor $\mathbb{E}_0$ is trivially an equivalence of categories.

@@

>>>>

\section*{Various issues-in-progress}
{\chris If Kenny likes, we can use this as a section for all the current issues that need to be resolved, or what we are currently working on or something!}
\begin{itemize}
\item {\chris I think structured cospans should be introduced/recalled way earlier! This is a paper that compares the two,
and both should be presented early.} Sure, I'm fine with that. What section do you think we should bring it up in, or maybe create a new section entirely to mention it? {\chris I think a new section right before that equivalence section would be great!}
\end{itemize}

@@

{\chris edit appropriately, notation etc.}
\begin{thm}\label{thm:Grothendieck}
    \begin{enumerate}
        \item Every opfibration $\U \maps \D \to \C$ gives rise to a pseudofunctor $F_U \maps \C {\to}Cat$.
        \item Every opindexed category $F \maps \C \to Cat$ gives rise to  an opfibration $\U_F \colon \int F\to\C$.
        \item The above correspondences yield an equivalence of 2-categories 
        \begin{displaymath}
            ICat(\C) \simeq opFib(\C)
        \end{displaymath}
        so that $F_{\U_F} \cong F$ and $\U_{F_\U} \cong \U$.
%        \item The above 2-equivalence extends to one between 2-categories of arbitrary-base fibrations and arbitrary-domain indexed categories
%        \begin{equation}\label{eq:Gr_equiv}
%        \ICat\simeq\Fib    
%        \end{equation}
    \end{enumerate}
\end{thm}

@@

{\chris  moved and motivated appropriately}
\begin{thm}\label{thm:mainthmDX}\cite{CV}
Suppose that $\D$ and $\C$ have chosen pushouts and initial objects, and a functor $\U\colon\D\to\C$ strictly preserves them. Then $\U$ is a right adjoint left inverse if and only if $\U$ is an opfibration.
\end{thm}

\section{An equivalence of symmetric monoidal double categories} \label{EquivDoubleCats}
In this section we prove that two different frameworks which utilize symmetric monoidal double categories, namely decorated cospans and structured cospans described in {\chris refer to somewhere}, are equivalent under the conditions required from the previous section that allow for a left adjoint $L \colon \mathsf{A} \to \mathsf{X}$ to be obtained from a symmetric lax {\chris have to specify lax VS weak which is the standard terminology, happy to use lax though} monoidal pseudofunctor $F \colon \mathsf{A} \to \mathbf{Cat}$. The double category version of decorated cospans is given in Section \ref{DecCospansDoubleCat}. The other framework also utilizing symmetric monoidal double categories due to the first two authors and which goes by the name of `structured cospans', is introduced below {\chris I see, prefer earlier!}. Once again following the notation of Shulman \cite{Shul2}, given a double category $\mathbb{A}$, we write $_f \mathbb{A}_g(M,N)$ for the set of 2-morphisms in $\mathbb{A}$ of the form:
\[
  \xymatrix@-.5pc{
    A \ar[r]|{|}^{M}  \ar[d]_f \ar@{}[dr]|{\Downarrow a}&
    B\ar[d]^g\\
    C \ar[r]|{|}_N & D
  }
\]
We call $M$ and $N$ the \define{horizontal source and target} of the 2-morphism $a$, respectively, and likewise we call $f$ and $g$ the \define{vertical source and target} of the 2-morphism $a$, respectively. Thus $_f \mathbb{A}_g(M,N)$ denotes the set of 2-morphisms in $\mathbb{A}$ with horizontal source and target $M$ and $N$ and vertical source and target $f$ and $g$.
\begin{defn}
A (possibly lax or oplax) double functor $\mathbb{F} \colon \mathbb{A} \to \mathbb{X}$ is \define{full} (respectively, \define{faithful}) if $\mathbb{F}_0 \colon \mathbb{A}_0 \to \mathbb{X}_0$ is full (respectively, faithful) and each map $$\mathbb{F}_1 \colon _f \mathbb{A}_g(M,N) \to _{\mathbb{F}(f)} \mathbb{X}_{\mathbb{F}(g)}(\mathbb{F}(M),\mathbb{F}(N))$$ is surjective (respectively, injective).
\end{defn}
\begin{defn}
A (possibly lax or oplax) double functor $\mathbb{F} \colon \mathbb{A} \to \mathbb{X}$ is \define{essentially surjective} if we can simultaneously make the following choices:
\begin{enumerate}
\item{For each object $x \in \mathbb{X}$, we can find an object $a \in \mathbb{A}$ together with a vertical 1-isomorphism $\alpha_x \colon \mathbb{F}(a) \to x$, and}
\item{For each horizontal 1-cell $N \colon x_1 \tobar x_2$  of $\mathbb{X}$, we can find a horizontal 1-cell $M \colon a_1 \tobar a_2$ of $\mathbb{A}$ and a 2-isomorphism $a_{N}$ of $\mathbb{X}$ as in the following diagram:
\[
  \xymatrix@-.5pc{
    \mathbb{F}(a_1) \ar[r]|{|}^{\mathbb{F}(M)}  \ar[d]_{\alpha_{x_1}} \ar@{}[dr]|{\Downarrow a_N}&
    \mathbb{F}(a_1) \ar[d]^{\alpha_{x_2}}\\
    x_1 \ar[r]|{|}_N & x_2
  }
\]
}
\end{enumerate}
\end{defn}
\begin{defn}
A double functor $\mathbb{F} \colon \mathbb{A} \to \mathbb{X}$ is \define{strong} if the comparison and unit constraints are globular isomorphisms, meaning that for each composable pair of horizontal 1-cells $M$ and $N$ we have a natural isomorphism $$\mathbb{F}_{M,N} \colon \mathbb{F}(M) \odot \mathbb{F}(N) \xrightarrow{\sim} \mathbb{F}(M \odot N)$$and for each object $a \in \mathbb{A}$ a natural isomorphism $$\mathbb{F}_a \colon \hat{U}_{\mathbb{F}(a)} \xrightarrow{\sim} \mathbb{F}(U_a).$$
\end{defn}
\begin{thm}[Shulman,7.8]\label{ShulDubEquiv}
Given a strong double functor $\mathbb{F} \colon \mathbb{A} \to \mathbb{X}$, $\mathbb{F}$ is part of a double equivalence if and only if $\mathbb{F}$ is full, faithful and essentially surjective.
\end{thm}

@@

{\chris in progress}
\begin{thm}\label{Equiv}
Let $\mathsf{A}$ and $\mathsf{X}$ be categories with (chosen) finite colimits, and $F \colon(\mathsf{A},+,0) \to (\mathbf{Cat},\times,\mathbf{1})$ a symmetric lax monoidal pseudofunctor. The symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$ of decorated cospans is equivalent to the symmetric monoidal double category $_L \mathbb{C}\textnormal{sp}(\mathsf{X})$ of structured cospans, where $L \colon \mathsf{A} \to \int{F} \cong \mathsf{X}$ is the left adjoint of the induced Grothendieck opfibration $\U_F\colon \int{F} \to \mathsf{A}$ by \cref{thm:Grothendieck} under the assumptions of \cref{thm:mainthmDX}.
\end{thm}
\begin{proof}
To prove this, we define a double functor $\mathbb{E} \colon _L \mathbb{C}\textnormal{sp}(\mathsf{X}) \to F\mathbb{C}\textnormal{sp}$ as follows: the object component of the double functor $\mathbb{E}$ is given by $\mathbb{E}_0 = \id_{\mathsf{A}}$ as both double categories $_L \mathbb{C}\textnormal{sp}(\mathrm{X})$ and $F\mathbb{C}\textnormal{sp}$ have objects and morphisms of $\mathsf{A}$ as objects and vertical 1-morphisms, respectively. The functor $\mathbb{E}_0$ is trivially an equivalence of categories.

@@

\bibitem{CV} D.\ Cicala and C.\ Vasilakopoulou, On Adjoints and Fibrations. In preparation.
